---
title: "Naive Bayes"
output: html_notebook
---
### Install required packages
```{r}
library(pacman)
p_load(ggplot2, tidyverse, caret, caretEnsemble, psych, 
       Amelia, mice, GGally, rpart, randomForest, e1071, rio)
```

### Load data
```{r}
sample <- import("./nbtrain.csv")
```

### Summary of sample
```{r}
summary(sample)
sample$income <- as.factor(sample$income)
sample$sex <- as.factor(sample$sex)
```

### Separate data on two sets for train and test model
```{r}
train_data <- as.data.frame(sample[1:9010,])
test_data <- as.data.frame(sample[9011:10010,])
```

# Naive Bayes Step 1
### Training model
```{r}
model <- naiveBayes(income ~ age + sex + educ, train_data)
model
```

### Testing model
```{r}
results <- predict(model, test_data)
results
confusion_matrix <- as.data.frame(table(results, test_data$income))
confusion_matrix
```

### Confusion matrix
```{r}
ggplot(data=confusion_matrix, mapping=aes(x=results, y=Var2)) +
  geom_tile(aes(fill=Freq)) +
  geom_text(aes(label=sprintf("%1.0f", Freq)), vjust=1) +
  scale_fill_gradient(low="blue", high="red", trans="log") +
  labs(y="Actual", x="Predicted")
```
### Missclassification rate
Another summary about sample 
```{r}
summary(test_data)
```

Overall
```{r}
miss_rate <- mean(results != test_data$income)
miss_rate
```

For low income
```{r}
miss_rate_low <- mean(results != test_data[test_data$income=="10-50K","income"])
miss_rate_low
```

For medium income
```{r}
miss_rate_medium <- mean(results != test_data[test_data$income=="50-80K","income"])
miss_rate_medium
```

For high income
```{r}
miss_rate_high <- mean(results != test_data[test_data$income=="GT 80K","income"])
miss_rate_high
```

# Naive Bayes Step 2
### Train model
```{r}
model_sex <- naiveBayes(sex ~ age + educ + income, train_data)
model_sex
```

### Testing model
```{r}
predicted_class <- predict(model_sex, test_data)
predicted_class
conf_matrix <- as.data.frame(table(predicted_class, test_data$sex))
conf_matrix
```

### Confusion matrix
```{r}
ggplot(data=conf_matrix, mapping=aes(x=predicted_class, y=Var2)) +
  geom_tile(aes(fill=Freq)) +
  geom_text(aes(label=sprintf("%1.0f", Freq)), vjust=1) +
  scale_fill_gradient(low="blue", high="red", trans="log") +
  labs(y="Actual", x="Predicted")
```

### Missclassification rate
Another summary about sample 
```{r}
summary(test_data)
```

Overall
```{r}
miss_rate_2 <- mean(predicted_class != test_data$sex)
miss_rate_2
```

For male
```{r}
miss_rate_male <- mean(predicted_class != test_data[test_data$sex=="M","sex"])
miss_rate_male
```

For female
```{r}
miss_rate_female <- mean(predicted_class != test_data[test_data$sex=="F","sex"])
miss_rate_female
```
