---
title: "Naive Bayes"
author: aeroshev
output: html_notebook
---
### Install required packages
```{r}
library(pacman)
p_load(ggplot2, tidyverse, caret, caretEnsemble, psych, 
       Amelia, mice, GGally, rpart, randomForest, e1071, rio)
```

### Load data
```{r}
sample <- import("./nbtrain.csv")
```

### Summary of sample
```{r}
summary(sample)
sample$income <- as.factor(sample$income)
sample$sex <- as.factor(sample$sex)
```

### Separate data on two sets for train and test model
```{r}
train_data <- as.data.frame(sample[1:9010,])
test_data <- as.data.frame(sample[9011:10010,])
```

# Naive Bayes Step 1
### Training model
```{r}
model <- naiveBayes(income ~ age + sex + educ, train_data)
model
```

### Testing model
```{r}
results <- predict(model, test_data)
results
confusion_matrix <- as.data.frame(table(results, test_data$income))
confusion_matrix
```

### Confusion matrix
```{r}
ggplot(data=confusion_matrix, mapping=aes(x=results, y=Var2)) +
  geom_tile(aes(fill=Freq)) +
  geom_text(aes(label=sprintf("%1.0f", Freq)), vjust=1) +
  scale_fill_gradient(low="blue", high="red", trans="log") +
  labs(y="Actual", x="Predicted")
```
### Missclassification rate
Another summary about sample 
```{r}
summary(test_data)
```

```{r}
confusion <- table(test_data$income, results)
confusion
```


Overall
```{r}
miss_rate <- mean(results != test_data$income)
miss_rate
```

For low income
```{r}
miss_rate_low <- 1 - (confusion[1, 1] / sum(confusion[1,]))
miss_rate_low
```

For medium income
```{r}
miss_rate_medium <- 1 - (confusion[2, 2] / sum(confusion[2,]))
miss_rate_medium
```

For high income
```{r}
miss_rate_high <- 1 - (confusion[3, 3] / sum(confusion[3,]))
miss_rate_high
```

# Naive Bayes Step 2
### Train model
```{r}
model_sex <- naiveBayes(sex ~ age + educ + income, train_data)
model_sex
```

### Testing model
```{r}
predicted_class <- predict(model_sex, test_data)
predicted_class
conf_matrix <- as.data.frame(table(predicted_class, test_data$sex))
conf_matrix
```

### Confusion matrix
```{r}
ggplot(data=conf_matrix, mapping=aes(x=predicted_class, y=Var2)) +
  geom_tile(aes(fill=Freq)) +
  geom_text(aes(label=sprintf("%1.0f", Freq)), vjust=1) +
  scale_fill_gradient(low="blue", high="red", trans="log") +
  labs(y="Actual", x="Predicted")
```

### Missclassification rate
Another summary about sample 
```{r}
summary(test_data)
```
```{r}
xconfusion <- table(test_data$sex, predicted_class)
xconfusion
```

Overall
```{r}
miss_rate_2 <- mean(predicted_class != test_data$sex)
miss_rate_2
```

For male
```{r}
miss_rate_male <- 1 - (xconfusion[2, 2] / sum(xconfusion[2,]))
miss_rate_male
```

For female
```{r}
miss_rate_female <- 1 - (xconfusion[1,1] / sum(xconfusion[1,]))
miss_rate_female
```

## Research
### Only male
```{r}
only_male <- train_data[train_data$sex=="M",]
only_male
```
### Only female
```{r}
only_female <- train_data[train_data$sex=="F",]
only_female
```
### Randomly get 3500 from female and male sets
```{r}
random_only_male <- sample_n(only_male, 3500)
random_only_female <- sample_n(only_female, 3500)
```

### Merge two frame
```{r}
random_data <- rbind(random_only_male, random_only_female)
random_data
```

### Train on random
```{r}
xmodel <- naiveBayes(sex ~ age + educ + income, random_data)
xmodel
```

```{r}
x_class <- predict(xmodel, test_data)
x_class
x_matrix <- as.data.frame(table(x_class, test_data$sex))
x_matrix
```

```{r}
ggplot(data=x_matrix, mapping=aes(x=x_class, y=Var2)) +
  geom_tile(aes(fill=Freq)) +
  geom_text(aes(label=sprintf("%1.0f", Freq)), vjust=1) +
  scale_fill_gradient(low="blue", high="red", trans="log") +
  labs(y="Actual", x="Predicted")
```

