---
title: "Clustering algorithms"
author: aeroshev
output: html_notebook
---
### Install packages
```{r}
library(pacman)
p_load(ggplot2, usmap, rio, magrittr, factoextra, NbClust, dplyr, cluster, fpc)
```


### Load data and set quantity of classes
```{r}
load("./income_elec_state.Rdata")
data <- as.data.frame(income_elec_state)
centers <- as.integer(4)
```

### That we have
```{r}
head(data)
```
### Setup seed for except random results for many runs
```{r}
set.seed(123)
```


### Clustering data by centers assigned up
```{r}
wss <- function(k) {
  kmeans(data, centers, nstart=10)$tot.withinss
}
points <- kmeans(data, centers, nstart=10)
```


### Join class number to data fram
```{r}
cluster_data <- cbind(data, cluster=points$cluster)
```

### After clustering
```{r}
head(cluster_data)
```


### Extract data about centroids of clusters
```{r}
centroids <- aggregate(formula=cbind(income, elec) ~ cluster, 
                       data=cluster_data, 
                       FUN=mean)
```


### Extract index states from data frame
```{r}
state <- row.names(cluster_data)
```


### Plot our data
```{r}
ggplot(data=cluster_data, aes(x=income, y=elec, colour=factor(cluster))) +
  geom_point(shape=1) +
  geom_text(aes(label=state), hjust=0, vjust=-0.5) +
  geom_point(data = centroids, size=4, shape=8) +
  labs(color="Clusters", x="Income", y="Elec", title="Plot of clustering data")
```

### Estimate our clusters
```{r}
fviz_nbclust(data, kmeans, method="wss")
```

## Clear data
### Remove PR and denomize income
```{r}
remove_these <- c("PR")
rows_to_remove <- which(row.names(data) %in% remove_these)

cdata <- data[-rows_to_remove,]
cdata$income <- apply(cdata, 1, function(x) x[1] / 70)

cpoints <- kmeans(cdata, centers, nstart=10)

ccluster_data <- cbind(cdata, cluster = cpoints$cluster)

centroids <- aggregate(formula=cbind(income, elec) ~ cluster, 
                       data=ccluster_data, 
                       FUN=mean)

cstate <- row.names(ccluster_data)

ggplot(data=ccluster_data, aes(x=income, y=elec, colour=factor(cluster))) +
  geom_point(shape=1) +
  geom_text(aes(label=cstate), hjust=0, vjust=-0.5) +
  geom_point(data = centroids, size=4, shape=8) +
  labs(color="Clusters", x="Income", y="Elec", title="Plot of clustering data")
```
### Estimate our clusters
```{r}
fviz_nbclust(cdata, kmeans, method="wss")
```



### US map for clear data
```{r}
state <- row.names(cdata)
usmap_data <- cbind(ccluster_data, state)
usmap_data$cluster <- as.factor(usmap_data$cluster)
plot_usmap(regions="states",
           data=as.data.frame(usmap_data),
           values="cluster",
           labels=TRUE) + 
  theme(legend.position="right") +
  labs(title="Clustering data by states")
```

## Experimant with value
```{r}
log_income <- log(data$income, base=10)
log_elec <- log(data$elec, base=10)

log_data <- as.data.frame(cbind(log_income, log_elec), row.names=state)

log_points <- kmeans(log_data, centers)

log_cluster_data <- cbind(log_data, cluster=log_points$cluster)

colnames(log_cluster_data) <- c("l_income", "l_elec", "l_cluster")

l_centroids <- aggregate(formula=cbind(l_income, l_elec) ~ l_cluster, 
                       data=log_cluster_data, 
                       FUN=mean)

head(log_cluster_data)
```


### Plot our data with modified data
```{r}
#ggplot(data=log_cluster_data,  inherit.aes=FALSE, aes(x=l_income, y=l_elec, #colour=factor(l_cluster))) +
#  geom_point(shape=1) +
#  geom_text(aes(label=state), hjust=0, vjust=-0.5) +
#  geom_point(data=l_centroids, size=4, shape=8) +
#  labs(color="Clusters", x="Income", y="Elec", title="Plot of clustering data")
```


# Hierarchical clustering
### Dissimilarity matrix
```{r}
gower.dist <- daisy(data, metric=c("gower"))
head(as.matrix(gower.dist))
```



```{r}
cstats.table <- function(dist, tree, k) {
clust.assess <- c("cluster.number","n","within.cluster.ss","average.within","average.between",
                  "wb.ratio","dunn2","avg.silwidth")
clust.size <- c("cluster.size")
stats.names <- c()
row.clust <- c()
output.stats <- matrix(ncol = k, nrow = length(clust.assess))
cluster.sizes <- matrix(ncol = k, nrow = k)
for(i in c(1:k)){
  row.clust[i] <- paste("Cluster-", i, " size")
}
for(i in c(2:k)){
  stats.names[i] <- paste("Test", i-1)
  
  for(j in seq_along(clust.assess)){
    output.stats[j, i] <- unlist(cluster.stats(d = dist, clustering = cutree(tree, k = i))[clust.assess])[j]
    
  }
  
  for(d in 1:k) {
    cluster.sizes[d, i] <- unlist(cluster.stats(d = dist, clustering = cutree(tree, k = i))[clust.size])[d]
    dim(cluster.sizes[d, i]) <- c(length(cluster.sizes[i]), 1)
    cluster.sizes[d, i]
    
  }
}
output.stats.df <- data.frame(output.stats)
cluster.sizes <- data.frame(cluster.sizes)
cluster.sizes[is.na(cluster.sizes)] <- 0
rows.all <- c(clust.assess, row.clust)
# rownames(output.stats.df) <- clust.assess
output <- rbind(output.stats.df, cluster.sizes)[ ,-1]
colnames(output) <- stats.names[2:k]
rownames(output) <- rows.all
is.num <- sapply(output, is.numeric)
output[is.num] <- lapply(output[is.num], round, 2)
output
}
```

### Division clustering from high to low
```{r}
divisive.clust <- diana(as.matrix(gower.dist),
                        diss=TRUE,
                        keep.diss=TRUE
                        )
plot(divisive.clust, main = "Divisive")
rect.hclust(aggl.clust.c, k=5, border="red")
```

### Elbow method for divisive
```{r}
ggplot(data = data.frame(t(cstats.table(gower.dist, divisive.clust, 15))), 
  aes(x=cluster.number, y=within.cluster.ss)) + 
  geom_point()+
  geom_line()+
  ggtitle("Divisive clustering") +
  labs(x = "Num.of clusters", y = "Within clusters sum of squares (SS)") +
  theme(plot.title = element_text(hjust = 0.5))
```


### Aglomerativ clustering for low to high
```{r}
aggl.clust.c <- hclust(gower.dist, method = "complete")
plot(aggl.clust.c,
     main = "Agglomerative, complete linkages")

rect.hclust(aggl.clust.c, k=6, border="red")
```

### Cluster Assessment
```{r}
ggplot(data = data.frame(t(cstats.table(gower.dist, aggl.clust.c, 15))), 
  aes(x=cluster.number, y=within.cluster.ss)) + 
  geom_point()+
  geom_line()+
  ggtitle("Agglomerative clustering") +
  labs(x = "Num.of clusters", y = "Within clusters sum of squares (SS)") +
  theme(plot.title = element_text(hjust = 0.5))
```
