---
title: "Association Rules"
author: aeroshev
output: html_notebook
---
### Installing Packages 
```{r}
library(pacman)
p_load(arules, arulesViz, ggplot2, tidyrules, dplyr, C50, pander)
```

### Read list of transactions
```{r}
dataset <- read.transactions("./AssociationRules.csv", sep=" ")
```

## Summary of dataset
### Max quantity
```{r}
freq_tab <- data.frame(itemFrequency(dataset, type="absolute"))
freq_tab <- cbind(rownames(freq_tab), freq_tab)
rownames(freq_tab) <- NULL
names(freq_tab) <- c("item", "freq")
freq_tab[freq_tab$freq == max(freq_tab$freq),]
```

### Most popular item
```{r}
itemFrequencyPlot(dataset, type="absolute", topN=10) 
```
```{r}
summary(dataset)
```


## Make association rules
```{r}
rules <- apriori(data=dataset, parameter=list(support=0.01, 
                                             confidence=0,
                                             target="rules",
                                             minlen=2))
```

### Summary rules
```{r}
summary(rules)
length(rules)
```
```{r}
inspect(rules)
```

### More confidence rules
```{r}
rules.hconf <- subset(rules, confidence >= 0.5)
length(rules.hconf)
```


```{r}
# Some 10 sample
rules.top <- sample(rules, 10)
# High confidence rules
rules.conf <- sort(rules, by="confidence", decreasing=TRUE)
# High lift rules
rules.hlift <- sort(rules, by="lift", decreasing=TRUE)
rules.llift <- sort(rules, by="lift", decreasing=FALSE)
# Top 10 rules by confidence
rules.top_conf <- head(rules.conf, n=10)
# Top 10 rules by lift
rules.top_lift = head(rules.hlift, n=10)
```


### Plot of all finded rules
```{r}
plot(rules, 
     method="scatterplot", 
     measure=c("support", "confidence"), 
     shading="lift",
     jitter=0.2
)
```

### Plot of hight confidence rules
```{r}
plot(rules.hconf, 
     method="scatterplot", 
     measure=c("support", "confidence"), 
     shading="lift",
     jitter=0.2
)
```

### Compare support and lift
```{r}
plot(rules, 
     method="scatterplot", 
     measure=c("support", "lift"), 
     shading="confidence",
     jitter=0.2
)
```
```{r}
plot(rules.hconf, 
     method="scatterplot", 
     measure=c("support", "lift"), 
     shading="confidence",
     jitter=0.2
)
```
```{r}
head(quality(rules.hconf))
```
```{r}
rules.hsup <- subset(rules, support >= 0.1)

plot(rules.hsup, 
     method="scatterplot", 
     measure=c("support", "confidence"), 
     shading="lift",
     engine="htmlwidget"
)
```

```{r}
inspect(head(rules.hsup, n=3, by="confidence", decreasong=TRUE))
```


### Point k
```{r}
inspect(head(rules.hlift, 10))
```
```{r}
plot(head(rules.hlift, 10), 
     method="matrix",
     measure="lift",
     control=list(recorder=FALSE),
     engine="htmlwidget")
```

```{r}
inspect(head(rules.llift, 10))
```
```{r}
plot(head(rules.llift, 10), 
     method="matrix", 
     measure="lift",
     control=list(recorder=FALSE),
     engine="htmlwidget")
```



```{r}
plot(rules, 
     method="scatterplot", 
     measure=c("support", "confidence"), 
     shading="lift",
     engine="htmlwidget"
)
```
```{r}
plot(rules, 
     method="scatterplot", 
     measure=c("support", "lift"), 
     shading="confidence",
     engine="htmlwidget"
)
```

```{r}
plot(rules.top_conf, method="paracoord")
```

```{r}
plot(rules.top_lift, method="paracoord")
```

```{r}
plot(rules.top_lift, method="grouped")
```

```{r}
plot(rules.top_conf, method="graph", engine="htmlwidget")
```
```{r}
top_3_lift = head(rules.top_lift, n=3)

plot(top_3_lift, method="graph")
```

```{r}
rules.hight_trustly <- subset(rules, confidence >= 0.8)

tail_40_conf = tail(sort(rules.hight_trustly, 
                          by="lift", 
                          decreasing=TRUE), n=40)

plot(rules.hight_trustly, 
     method="matrix", 
     shading=c("lift", "confidence"), 
     measure=c("lift", "confidence"), 
     control=list(reorder=FALSE)
)
```
## Train model prediction
```{r}
dataset.train <- dataset[1:8000,]
dataset.test <- dataset[8001:10000,]
```

```{r}
model <- apriori(data=dataset.train, parameter=list(support=0.01, 
                                                    confidence=0.1,
                                                    target="rules",
                                                    minlen=2))
validate_model <- apriori(data=dataset.test, parameter=list(support=0.01,
                                                            confidence=0.1,
                                                            target="rules",
                                                            minlen=2))
```

```{r}
train_df <- DATAFRAME(model)
validate_df <- DATAFRAME(validate_model)
```

### Mean support for trained rules 
```{r}
cat('Support train:\t', mean(train_df$support), '\n')
```

### Mean confidence for trained rules 
```{r}
cat('Confidence train:\t', mean(train_df$confidence), '\n')
```

### Mean support for validate rules 
```{r}
cat('Support validate:\t', mean(validate_df$support), '\n')
```

### Mean confidence for validate rules 
```{r}
cat('Confidence validate:\t', mean(validate_df$confidence), '\n')
```

```{r}

```


