---
title: "R Notebook"
output: html_notebook
---
I install package in UI R Studio, but you can install with commands
```{r}
install.packages("pacman")
library(pacman)
```
Import require libraries wit help pacman
```{r}
pacman::p_load(pacman, dplyr, GGally, ggplot2, ggthemes, ggvis, httr, lubridate,
plotly, rio, rmarkdown, shiny, stringr, tidyr)
```
Import data from file
```{r}
data <- import("./Data/zipIncome.txt")
```
Rename columns
```{r}
colnames(data) <- c("zipCode", "income")
```
What we have
```{r}
head(data)
```
Summary of our data
```{r}
summary(data)
```
Plot of our data
```{r}
plot(data)
```
Same plot, but use ggplot
```{r}
ggplot(data, aes(x = zipCode, y = income)) + 
  geom_point()
```
Clear our data in range income 7.000 $ < income < 200.000 $ 
And in same time rename columns, because split function return list and we should 
transform to dataframe
```{r}
clear_data <- as.data.frame(split(data, cut(data$income, c(7000, 200000))))
colnames(clear_data) <- c("zipCode", "income")
```
That we gave now
```{r}
head(clear_data)
```
New summary of clear data
```{r}
summary(clear_data)
```
This is boxplot of our clearing data
```{r}
boxplot(clear_data$income ~ clear_data$zipCode, 
        ylab = "Income", 
        xlab = "Zip code", 
        main = "Plot of income by zip code")
```
Plotted out y scale in log scale
```{r}
boxplot(clear_data$income ~ clear_data$zipCode,
        log = "y",
        ylab = "Income", 
        xlab = "Zip code", 
        main = "Plot of income by zip code with log dependicies")
```
Set zip code as factor for grouping data by zip code
```{r}
clear_data$zipCode <- as.factor(clear_data$zipCode)
```
Use ggplot for groupping income by zip code
```{r}
ggplot(clear_data, aes(x = zipCode, y = income)) +
  ylab("Income") +
  geom_jitter(alpha = 0.2) + 
  scale_y_log10()
```
Add box plot to current plot new layout
```{r}
ggplot(clear_data, aes(x = as.factor(zipCode), y = income)) +
  geom_jitter(aes(col = factor(zipCode)), alpha = 0.2) +
  geom_boxplot(alpha = 0.1, outlier.size = -Inf) + 
  scale_y_log10() +
  labs(color="Region", x="Zip Code", y="Income", title="Average Income by Zip Code")
```