---
title: "Linear regression"
author: "aeroshev"
output: html_notebook
---
### Load modules
```{r}
library(pacman)
p_load(rio, dplyr, hexbin, RColorBrewer, ggplot2)
```

# Load anf prepair data
### Load data
```{r}
df <- as.data.frame(import("./zeta.csv"))
df <- transform(df, 
          meanage=as.numeric(meanage), 
          meaneducation=as.numeric(meaneducation), 
          meanhouseholdincome=as.numeric(meanhouseholdincome))
head(df)
```

### Remove duplicates "meanhouseholdincome"
```{r}
df <- df %>% distinct(meanhouseholdincome, .keep_all=TRUE)
head(df)
```

### Remove columns "zcta" and "sex"
```{r}
df <- within(df, rm(V1, zcta, sex))
head(df)
```
### Remove outliers
```{r}
df <- df %>% filter(between(meaneducation, 8, 18)) %>%
             filter(between(meanhouseholdincome, 10000, 200000)) %>%
             filter(between(meanemployment, 0, 3)) %>%
             filter(between(meanage, 20, 60))
head(df)
```

### Create a new variable 
```{r}
df <- cbind(df, log10(df$meanhouseholdincome))
head(df)
```

### Reaname columns
```{r}
colnames(df) <- c("age", "education", "employment", "income", "log_income")
head(df)
```

# Analysis
`a.` ##  We will be analyzing this data with income as the dependent variable 
and the other columns as independent variables.  
Create a scatter plot showing the effect age has on log_income and paste it here.  
Do you see any linear relationship between the two variables? <br>
### Simple plot
```{r}
plot(df$log_income ~ df$age, 
     ylab="log_income", 
     xlab="age",
     main="Primary log income ~ age",)
with(df, lines(loess.smooth(age, log_income), col="red"))
```
### Pretty plot
```{r}
opacity <- colorRampPalette(rev(brewer.pal(11,'Spectral')))
hexbinplot(df$log_income ~ df$age, 
           ylab="log_income", 
           xlab="age",
           colramp=opacity)
```
### Plot with prediction
```{r}
hexbinplot(df$log_income ~ df$age, 
           trans=sqrt, 
           inv=function(x) x^2, 
           type=c("r"),
           ylab="log_income",
           xlab="age",
           main="Regression of log income ~ age",
           colramp=opacity)
```

### Create a linear regression <br>
`b.` 
Create a linear regression model between log_income and age. 
What is the interpretation of the t-value? 
What kind of t-value would indicate a significant coefficient?
```{r}
model_age <- lm(formula=log_income ~ age, data=df)
summary(model_age)
```
```{r}
cor(df$log_income, df$age)
```

`c.` What is the interpretation of the R-squared value?  
What kind of R-squared value would indicate a good fit?
`d.` What is the interpretation of the F-statistic?  
What kind of F-statistic indicates a strong linear regression model?
`e.` View a detailed summary of the previous model.
What is the R-squared value?  
Does this suggest that the model is a good fit? Why?
`f.` Create a scatter plot showing the effect education has on log_income.  
Do you see any linear relationship between the two variables?
```{r}
plot(df$log_income ~ df$education, 
     ylab="log_income", 
     xlab="education",
     main="Primary log income ~ education")
with(df, lines(loess.smooth(education, log_income), col="red"))
```
### Pretty plot
```{r}
hexbinplot(df$log_income ~ df$education, 
           ylab="log_income", 
           xlab="education", 
           type=c("r"),
           main="Regression of log income ~ education",
           colramp=opacity)
```

`g.` Analyze a detailed summary of a linear regression model between log_income and education.  What is the R-squared value?  Is the model a good fit?  Is it better than the previous model?
```{r}
model_education <- lm(formula=log_income ~ education, data=df)
summary(model_education)
```

`h.` Analyze a detailed summary of a linear regression model between the dependent variable log_income, and the independent variables age, education, and employment.  Is this model a good fit?  Why?  What conclusions can be made about the different independent variables?
```{r}
model <- lm(formula=log_income ~ age + education + employment, data=df)
summary(model)
```
```{r}
df$predict <- predict(model, select(df, "age", "education", "employment"))
plot(df$predict ~ df$log_income, 
     ylab="Predicted data", 
     xlab="Training data",
     main="Analyze results model")
with(df, lines(loess.smooth(log_income, predict), col="blue"))
abline(a=0, b=1, col="red")
```

### Pretty plot
```{r}
hexbinplot(df$predict ~ df$log_income, 
           ylab="Predicted Income", 
           xlab="Training Income", 
           trans=sqrt, 
           inv=function(x) x^2, 
           type=c("r"),
           main="Analyze results model",
           panel=function(x, y, ...) {
             panel.hexbinplot(x, y, ...)
             lattice::panel.abline(a = 0, b = 1, col="red")
           },
           colramp=opacity)
```
### Trend of age + education + employment for linear model
```{r}
for (i in 1:nrow(df)) {
  df[i, "age_education_employment"] <- sum(df[i, "age"], 
                                          df[i, "education"], 
                                          df[i, "employment"])
}
hexbinplot(df$log_income ~ df$age_education_employment,
           ylab="Log income", 
           xlab="age + education + employment", 
           trans=sqrt, 
           inv=function(x) x^2, 
           type=c("r"),
           main="Regression of log_income ~ age + education + employment",
           colramp=opacity)
```

### Primary data trend
```{r}
plot(log_income ~ age_education_employment, 
     data=df,
     xlab="age + education + employment",
     ylab="log income",
     main="Primary log income ~ age + education + employment", 
     col="black"
)
with(df, lines(loess.smooth(age_education_employment, log_income), col="red"))
```
### Corelasion between
```{r}
cor(df$log_income, df$age_education_employment)
```

### Try create more interesting
```{r}
best_model <- lm(formula=log_income ~ education + employment, data=df)
summary(best_model)
```
### Primary data
```{r}
for (i in 1:nrow(df)) {
  df[i, "education_employment"] <- sum(df[i, "education"],
                                       df[i, "employment"])
}
plot(log_income ~ education_employment, 
     data=df,
     xlab="education + employment",
     ylab="log income",
     main="Primary log income ~ education + employment", 
     col="black"
)
with(df, lines(loess.smooth(education_employment, log_income), col="red"))
```
### Regression of log income ~ education + employment
```{r}
hexbinplot(df$log_income ~ df$education_employment,
           ylab="Log income", 
           xlab="education + employment", 
           trans=sqrt, 
           inv=function(x) x^2, 
           type=c("r"),
           main="Regression of log_income ~ education + employment",
           colramp=opacity)
```
### Accuracy model
```{r}
df$best_predict <- predict(best_model, select(df, "education", "employment"))

hexbinplot(df$best_predict ~ df$log_income, 
           ylab="Predicted Income", 
           xlab="Training Income", 
           trans=sqrt, 
           inv=function(x) x^2, 
           type=c("r"),
           main="Analyze results model",
           panel=function(x, y, ...) {
             panel.hexbinplot(x, y, ...)
             lattice::panel.abline(a = 0, b = 1, col="red")
           },
           colramp=opacity)
```
```{r}
plot(df$best_predict ~ df$log_income, 
     ylab="Predicted data", 
     xlab="Training data",
     main="Analyze results model")
with(df, lines(loess.smooth(log_income, best_predict), col="blue"))
abline(a=0, b=1, col="red")
```

```{r}
AIC(model)
```
```{r}
AIC(best_model)
```
```{r}
?hexbinplot
```

